---
title: "In Class Exercise 5"
author: "Hao Xian"
date: "6 February 2023"
date-modified: '`r Sys.Date()`'
format: html
number-sections: true
execute: 
  echo: true
  eval: true
  message: false
  warning: false
  
editor: visual
---

# Getting Started

```{r}
pacman::p_load(tidyverse, tmap, sf, sfdep)
```

# Importing Data

::: callout-note
Taiwan has 2 projection system
:::

```{r}
studyArea <- st_read(dsn = "data",
                     layer="study_area") %>%
  st_transform (crs =3829)
```

```{r}
stores <- st_read(dsn = "data",
                     layer="stores") %>%
  st_transform (crs =3829)
```

```{r}
tmap_mode("view")
tm_shape(studyArea)+
  tm_polygons() +
  tm_shape(stores) +
  tm_dots(col="Name",
          size = 0.01,
          border.col="black",
          border.lwd=0.5)+
  tm_view(set.zoom.limits = c(12,16))
```

# Local Colocation Quotients (LCLQ)

::: callout-note
6 nearest neighbour is used. Include Self is needed as it help to boost the number to an odd number.

This is an adaptive method. Weights of the stores will be stored when st_knernel_weights is used
:::

```{r}
nb <- include_self(
  st_knn(st_geometry(stores),6)
)
```

```{r}
wt <- st_kernel_weights(nb,
                        stores,
                        "gaussian",
                        adaptive = TRUE)
```

```{r}
FamilyMart <- stores %>%
  filter(Name == "Family Mart")
A <- FamilyMart$Name
```

```{r}
SevenEleven <- stores %>%
  filter(Name == "7-Eleven")
B <- SevenEleven$Name
```

::: callout-note
A is the target

B is the neightbour

nb is a list of nearest neighbour

wt is the weight

49 is the simulation runs

The simulation will run 50 Times, and will generate a different p-value

NA is common in local_colocation()
:::

```{r}
LCLQ <- local_colocation (A, B, nb, wt, 49)
```

::: callout-note
There is no unique idenitfier therefore, the only code is cbind. cbind cannot work if the results is sorted.
:::

```{r}
LCLQ_stores <- cbind(stores, LCLQ)
```

```{r}
tmap_mode("view")+
  tm_shape(studyArea)+
  tm_polygons() +
  tm_shape(LCLQ_stores) +
  tm_dots(col="X7.Eleven",
          size = 0.01,
          border.col="black",
          border.lwd=0.5)+
  tm_shape(stores)+
  tm_dots(col="Name",
          size = 0.01,
          border.col="black",
          border.lwd=0.5)
  tm_view(set.zoom.limits = c(12,16))
```
