---
title: "In Class Exercise 8"
author: "Hao Xian"
date: "6 March 2023"
date-modified: '`r Sys.Date()`'
format: html
number-sections: true
execute: 
  echo: true
  eval: true
  message: false
  warning: false
  
editor: visual
---

::: callout-important
We will be using sfdep in In-Class-Excercise instead of spded in Hands On Excercise
:::

# Loading the packages

::: callout-note
Remove Corrplot, as there might be some conflicts
:::

```{r}

pacman::p_load(olsrr, sf, spdep, GWmodel, tmap, gtsummary, ggplot)
```

# The data

## Importing the Geospatial data

::: callout-note
This file is loaded as a sf dataframe
:::

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

## Update the CRS Model

```{r}
st_crs(mpsz_svy21)
```

```{r}
st_bbox(mpsz) #view extent
```

## Reading the Aspatial Data

```{r}
condo_resale = readr::read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
dplyr::glimpse(condo_resale)
```

```{r}
summary(condo_resale)
```

### Converting aspatial dataframe into a sf object

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

# EDA

```{r}
## Need to fix
##ggplot::ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
##  geom_histogram(bins=20, color="black", fill="light blue")
```

Since the data is skewed we should normalize it by logging it

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

```{r}
## need to fix
##ggplot::ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
##  geom_histogram(bins=20, color="black", fill="light blue")
```

The distribution is now less skewed

## Multiple Histogram Plots

```{r}

## Need to fix all of t


```

### Drawing Statistical Point Map

```{r}
tmap_mode("view")
```

```{r}
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tmap_options(check.and.fix = TRUE)+
  tm_dots(
          col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

# Building Hedonic Price Model

## Simple Linear Regression

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
```

We will plot it to see

```{r}
## Need to fix thios

```

## Multiple Linear Regression

We need to check the correlation first

```{r}
corrplot::corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

## Building a hedonic pricing model using multiple linear regression method

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

### Preparing Publication Quality Table

```{r}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                 data=condo_resale.sf)
ols_regress(condo.mlr1)
```

### Preparing using GT SUmamry

```{r}
tbl_regression(condo.mlr1, intercept = TRUE)

```

### Regression

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

### CHecking for multicolinearity

```{r}
ols_vif_tol(condo.mlr1)
```
