---
title: "In Class Exercise 4"
author: "Hao Xian"
date: "30 January 2023"
date-modified: '`r Sys.Date()`'
format: html
number-sections: true
execute: 
  echo: true
  eval: true
  message: false
  warning: false
  
editor: visual
---

# Loading the package

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

Things to learn outside from this code chunk.

# Load Dataset

```{r}
childcare_sf <- st_read("data/geospatial/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

This is already a projected coordinate system. This is converted to meters.

::: callout-note
For all spatial analysis, all data must be projected coordinate system.
:::

```{r}
sg_sf <- st_read(dsn = "data/geospatial", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

### Converting to Spatial Dataframe

::: callout-note
Spatial Polygon has a data table
:::

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

### Converting to Spatial Object

::: callout-note
This only retains the geometric
:::

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

### Converting to PPP format

::: callout-note
This drops everything else into a list of coordinates that can only be read by the ST package
:::

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```
